### 2020-12-25 Modules/arrayPerformance_beamwidth.tex
Before applying meeting#1's comments

The half power beamwidth (HPBW), denoted as $\dThetaHPBW$ quantifies the array's main lobe narrowness.
It marks the DOA where the beampattern's energy reduces to half of its maximal value.
For standard ULA, assuming large $N$ values, it is known \cite{van2004optimum} that
$$\dThetaHPBW/2= 1.4/N.$$
\par In App.~\ref{apdx_HPBW} we extend this known result for any $r\geq 0$. It turns out that for large $N$, the HPBW is obtained by solving for $x$ the equality
\begin{equation}\label{eq_HPBW}
    % \resizebox{1\linewidth}{!}{
        \begin{split}
            \rBrace{r^{2}-4r+2}\frac{\sin{\rBrace{x}}^{2}}{x^{2}}+r\frac{\sin{\rBrace{2x}}}{x}-1=0
        \end{split}
    % }
\end{equation}
where we define $x\triangleq{} N\dTheta_{HPBW}/2$. In Fig.~\ref{fig_feedbackULA_HPBW_Nx_vs_N_variousR} we plot the numerical solution of \eqref{eq_HPBW} for various values of $r$ and $N$, showing that $x$ reaches its limit around $N=20$. Also note that for $r=0$ we obtain the known result of standard ULA with the limiting factor of $1.4$.
Having the limiting factors for various values of the gain mismatch $r$, we investigate the feedback related improvement and express the HPBW by
\[
\dThetaHPBW/2\approx \frac{1.4}{f(r)N}.
\]
Note that $f(r)$ represents the array aperture improvement factor, compared to the standard ULA. 
Numerical evaluation (see Fig.~\ref{fig_feedbackULA_beamwidth_limit_r_dependent}) shows that $f(r)$ may be approximated with a second order polynomial
\begin{equation}
    \label{eq_Bapprox}
    f\rBrace{r}\approx\frac{1.4}{\rBrace{1-r}\rBrace{-0.4r+1.4}}.
\end{equation}
Note that for accurate gain match (i.e., $r\to1$), the RHS of \eqref{eq_Bapprox} tends towards infinity, implying that the equivalent array has an infinite number  of elements ($f\rBrace{r}N$), hence obtaining perfect spatial selectivity.
\begin{figure}[t]
    \begin{center}
        \begin{overpic}[width=0.65\linewidth, 
        %grid, 
        tics=10,trim=0 0 0 0]{./Media/spatial_IIR_MATLAB/arrayParameters/HPBW_vs_N_various_r.eps}
            \put (4, 75){\footnotesize{$N\dThetaHPBW/2$}}
            \put (50, 62.5) {\footnotesize{$r=0$}}
            \put (50, 54) {\footnotesize{$r=0.1$}}
            \put (50, 39.5) {\footnotesize{$r=0.3$}}
            \put (50, 28.5) {\footnotesize{$r=0.5$}}
            \put (50, 19.75) {\footnotesize{$r=0.7$}}
            \put (50, 12.5) {\footnotesize{$r=0.9$}}
            \put (50, 2) {\footnotesize{$N$}}
        \end{overpic}
    \end{center}
     \caption{Plot of $x=N\dThetaHPBW/2$ vs. $N$, for various $r$ values, obtained by numerically solving \eqref{eq_HPBW}.}
    \label{fig_feedbackULA_HPBW_Nx_vs_N_variousR}
\end{figure}
\begin{figure}[t]
    \begin{center}
        \begin{overpic}[width=0.65\linewidth, 
        % grid, 
        tics=10,trim=0 0 0 0]{./Media/HPBW_limit_vs_r.eps}
            \put (39.5, 63.5) {\scriptsize{Numerical solution of \eqref{eq_HPBW}}}
            \put (39.5, 58.25) {\scriptsize{Polynomial fitting \eqref{eq_Bapprox}}}
            \put (39.5, 52.5) {\footnotesize{$\log_{10}f\rBrace{r}$}}
            \put (85, 75) {\footnotesize{$\log_{10}f\rBrace{r}$}}
            \put (4, 75){\footnotesize{$N\dThetaHPBW/2$}}
            \put (50, 2) {\footnotesize{$r$}}
        \end{overpic}
    \end{center}
    \caption{Evaluation of $N\dThetaHPBW/2$ for $N=100$ and its approximation $1.4/f\rBrace{r}$  (marked by red diamonds). $f\rBrace{r}$ is also presented, in logarithmic scale (dotted curve).} 
    \label{fig_feedbackULA_beamwidth_limit_r_dependent}
\end{figure}

>>>>

The half power beamwidth (HPBW), denoted as $\dThetaHPBW$ quantifies the array's main lobe narrowness.
It represents the DOA where the beampattern's energy reduces to half of its maximal value.
For standard ULA, assuming large $N$ values, it is known \cite{van2004optimum} that
$$\dThetaHPBW/2= 1.4/N.$$
\par In App.~\ref{apdx_HPBW} we extend this known result for any $r\geq 0$. It turns out that for large $N$, the HPBW is obtained by solving for $x$ the equality
\begin{equation}\label{eq_HPBW}
    % \resizebox{1\linewidth}{!}{
        \begin{split}
            \rBrace{r^{2}-4r+2}\frac{\sin{\rBrace{x}}^{2}}{x^{2}}+r\frac{\sin{\rBrace{2x}}}{x}-1=0
        \end{split}
    % }
\end{equation}
where we define $x\triangleq{} N\dTheta_{HPBW}/2$. In Fig.~\ref{fig_feedbackULA_HPBW_Nx_vs_N_variousR} we plot the numerical solution of \eqref{eq_HPBW} for various values of $r$ and $N$, showing that $x$ reaches its limit around $N=20$. Also note that for $r=0$ we obtain the known result of standard ULA with the limiting factor of $1.4$.
Having the limiting factors for various values of the gain mismatch $r$, we investigate the feedback related improvement and express the HPBW by
\[
\dThetaHPBW/2\approx \frac{1.4}{f(r)N}.
\]
Note that $f(r)$ represents the array aperture improvement factor, compared to the standard ULA. 
Numerical evaluation (see Fig.~\ref{fig_feedbackULA_beamwidth_limit_r_dependent}) shows that $f(r)$ may be approximated with a second order polynomial
\begin{equation}
    \label{eq_Bapprox}
    f\rBrace{r}\approx\frac{1.4}{\rBrace{1-r}\rBrace{-0.4r+1.4}}.
\end{equation}
Note that for an accurate gain match (i.e., $r\to1$), the RHS of \eqref{eq_Bapprox} tends towards infinity, implying that the equivalent array has an infinite number  of elements ($f\rBrace{r}N$), hence obtaining perfect spatial selectivity.
\begin{figure}[t]
    \begin{center}
        \begin{overpic}[width=0.65\linewidth, 
        %grid, 
        tics=10,trim=0 0 0 0]{./Media/spatial_IIR_MATLAB/arrayParameters/HPBW_vs_N_various_r.eps}
            \put (4, 75){\footnotesize{$N\dThetaHPBW/2$}}
            \put (50, 62.5) {\footnotesize{$r=0$}}
            \put (50, 54) {\footnotesize{$r=0.1$}}
            \put (50, 39.5) {\footnotesize{$r=0.3$}}
            \put (50, 28.5) {\footnotesize{$r=0.5$}}
            \put (50, 19.75) {\footnotesize{$r=0.7$}}
            \put (50, 12.5) {\footnotesize{$r=0.9$}}
            \put (50, 2) {\footnotesize{$N$}}
        \end{overpic}
    \end{center}
     \caption{Plot of $x=N\dThetaHPBW/2$ vs. $N$, for various $r$ values, obtained by numerically solving \eqref{eq_HPBW}.}
    \label{fig_feedbackULA_HPBW_Nx_vs_N_variousR}
\end{figure}
\begin{figure}[t]
    \begin{center}
        \begin{overpic}[width=0.65\linewidth, 
        % grid, 
        tics=10,trim=0 0 0 0]{./Media/HPBW_limit_vs_r.eps}
            \put (39.5, 63.5) {\scriptsize{Numerical solution of \eqref{eq_HPBW}}}
            \put (39.5, 58.25) {\scriptsize{Polynomial fitting \eqref{eq_Bapprox}}}
            \put (39.5, 52.5) {\footnotesize{$\log_{10}f\rBrace{r}$}}
            \put (85, 75) {\footnotesize{$\log_{10}f\rBrace{r}$}}
            \put (4, 75){\footnotesize{$N\dThetaHPBW/2$}}
            \put (50, 2) {\footnotesize{$r$}}
        \end{overpic}
    \end{center}
    \caption{Evaluation of $N\dThetaHPBW/2$ for $N=100$ and its approximation $1.4/f\rBrace{r}$  (marked by red diamonds). $f\rBrace{r}$ is also presented, in logarithmic scale (dotted curve).} 
    \label{fig_feedbackULA_beamwidth_limit_r_dependent}
\end{figure}

### 2020-12-25 Modules/estimation_CRLB_appendix.tex
Before applying meeting#1's comments

As the feedback-based transfer function (\ref{eqn:GeneralFeedbackTransferFunction}) is expressed in frequency domain, we rely on \cite{zeira1990frequency} to express the frequency domain FIM as well. 

@@

For simplicity, $\text{n}\rBrace{t}$ is assumed to be a white Gaussian with some constant power spectral density $\Phi(\omega)=\sigma^2$ and independent of the estimated parameters $\vEta$. Hence, the second term in the right-hand-side of \eqref{eq_FIM_kl_full} vanishes. 

@@

\frac{\partial\vd}{\partial\thetaD}=\vecnot{A}\omegaB\vd

@@

where $\vecnot{A}\omegaB$ is an $N\times{}N$ diagonal matrix with

@@

A_{ii}\omegaB=-j\omega\frac{\partial \tau_{i}}{\partial{\thetaD}}\ \  \forall{i\in\cBrace{0\hdots{}N-1}}.

@@

            \lBrace{g\vBetaH{}\vecnot{A}\omegaB\vd-g^{2}\vBetaH{}B\omegaB\vAlphaC\ePhi{-}}^{2}

@@

where $\vecnot{B}\omegaB\triangleq\vd\vdT{}\vecnot{A}\omegaB-\vecnot{A}\omegaB\vd\vdT.$ 

>>>>

As the feedback-based transfer function (\ref{eqn:GeneralFeedbackTransferFunction}) is expressed in the frequency domain, we rely on \cite{zeira1990frequency} to express the frequency domain FIM as well. 

@@

For simplicity, $\text{n}\rBrace{t}$ is assumed to be white Gaussian with some constant power spectral density $\Phi(\omega)=\sigma^2$ and independent of the estimated parameters $\vEta$. Hence, the second term in the RHS of \eqref{eq_FIM_kl_full} vanishes. 

@@

\frac{\partial\vd}{\partial\thetaD}=\vecnot{A}\vd

@@

where $\vecnot{A}$ is an $N\times{}N$ diagonal matrix with

@@

A_{ii}=-j\omega\frac{\partial \tau_{i}}{\partial{\thetaD}}\ \  \forall{i\in\cBrace{0\hdots{}N-1}}.

@@

            \lBrace{g\vBetaH{}\vecnot{A}\vd-g^{2}\vBetaH{}B\vAlphaC\ePhi{-}}^{2}

@@

where $\vecnot{B}\triangleq\vd\vdT{}\vecnot{A}-\vecnot{A}\vd\vdT.$ 

### 2020-12-25 Modules/sec_conclusions.tex
Before applying meeting#1's comments

\section{Conclusions}
\label{sec_conclusions}
Integrating feedback into standard beamformers proved to achieve the spatial domain equivalent of the temporal IIR filtering.
It seems that a simple generalization of the conventional-beamformer maximizes (locally) the system's residing spatial information, thus enabling high localization accuracy.
The feedback-based architecture performance evaluation predicts an unlimited improvement in all criteria, when considering perfect knowledge of the target's range and the channel attenuation.
It turns out that a single frequency waveform based feedback-beamformer is impractical, being too sensitive to even mild target range estimation errors.
Fortunately, using two frequency waveform and applying simple frequency domain manipulations to the output and feedback signals, were found to serve as a low frequency (hence low sensitivity) equivalent of the single frequency scheme.
Also, the dual frequency scheme proved to be of low noise sensitivity, featuring high performance even in relatively low signal-to-noise-ratio scenarios.
\par Future study of the feedback beamforming concept may be applied to other array processing applications other than localization.
Furthermore, it is worthwhile to inspect other interesting choices of coefficients rather than the conventional-beamformer generalization, other waveforms and associated processing schemes, to extend the results to dynamic/multiple targets and to consider sensors with general radiation patterns.
Also, one may consider generalizing the suggested architecture to multiple-input-multiple-output systems, enabling a steered/focused (rather than omni-directional) feedback transmission.

>>>>

Integrating feedback into standard beamformers proved to achieve the spatial domain equivalent of the temporal IIR filtering.
It seems that a simple generalization of the conventional-beamformer maximizes (locally) the system's spatial information, thus enabling high localization accuracy.
The feedback-based architecture performance evaluation predicts an unlimited improvement in all criteria, when considering perfect knowledge of the target's range and the channel attenuation.
It turns out that a single frequency waveform based feedback-beamformer is impractical, being too sensitive to even mild target range estimation errors.
Fortunately, using a dual-frequency waveform and applying simple frequency domain manipulations to the output and feedback signals, were found to serve as a low frequency (hence low sensitivity) equivalent of the single frequency scheme.
Also, the dual frequency scheme proved to be of low noise sensitivity, featuring high performance even in relatively low signal-to-noise-ratio scenarios.
\par Future study of the feedback beamforming concept may be applied to other array processing applications other than localization.
Furthermore, it is worthwhile to inspect other interesting choices of coefficients rather than the conventional-beamformer generalization, other waveforms and associated processing schemes, to extend the results to dynamic/multiple targets and to consider sensors with general radiation patterns.
Also, one may consider generalizing the suggested architecture to multiple-input-multiple-output systems, enabling a steered/focused (rather than omni-directional) feedback transmission.

### 2020-12-25 Modules/sec_prlm_propWaveField.tex
Before applying meeting#1's comments

In array signal processing, propagating waves carry signals from the source to the array.
Therefore, these signals are represented in the time-space domain. The space domain is represented by either the three dimensional Cartesian coordinates $\rBrace{x,y,z}$ or the three dimensional spherical coordinates $\rBrace{r,\phi,\theta}$, where $0 \leq \phi \leq 2\pi$, $0 \leq \theta \leq \pi$ are the azimuth and elevation angles, respectively. The time domain is represented by $t$.
Finally, let $f\rBrace{t,\vecnot{r}}$ denote the time-space representation of the input signal, where $\vecnot{r}$ is the radius-vector in the three-dimensional system. The relations between the Cartesian and spherical coordinates are given in Fig.~\ref{fig_coordinates}.
$f\rBrace{t,x,y,z}$ describes the signal impinging the microphone array. The physics of propagation within a homogeneous (constant speed of propagation), dispersion free (no signal degradation due to other frequencies) and lossless (no signal gain attenuation) medium is described by the wave equation:

@@

Where $\nabla^{2}$ represents the Laplacian operator and $c$ represents the wave's velocity.

@@

Let $f_p\rBrace{t,x,y,z}$ denote a possible solution to \eqref{eq_prlm_waveEq} of a complex exponential form:

@@

Substituting \eqref{eq_prlm_waveEq_pSol} into \eqref{eq_prlm_waveEq} yields:

@@

This solution is called a monochromatic plane wave. The term plane wave is derived from the fact that at any given time, $t_{0}$, the value of $f\rBrace{t_{0},x,y,z}$ is the same at all points on a plane given by $k_{x}x+k_{y}y+k_{z}z = constant$.
Finally, writing the solution to \eqref{eq_prlm_waveEq} using vector notation we obtain:

@@

Let $\vecnot{k}\vecnot{x}=constant$ denote planes of constant phase. 
If the signal $f\rBrace{t,\vecnot{x}}$ is indeed propagating, planes of constant phase move by small steps of $\delta{\vecnot{x}}$, with small time increments of $\delta{t}$, i.e. $f\rBrace{t+\delta{t},\vecnot{x}+\delta{\vecnot{x}}} = f\rBrace{t,\vecnot{x}}$, which yields

@@

Assuming that $c>0$.
Thus, the speed of propagation is indeed $c$.
Let $\lambda$ denote the distance the plane wave propagated during a single temporal period of $T=\frac{2\pi}{\omega}$.
This distance is referred to as the \emph{wavelength}. 
Due to its spatial traits, after one period the monochromatic plane wave appears the same as one period before, but moved forward.

@@

Using \eqref{eq_prlm_waveEq_stepsEq} with $delta{t} = \frac{2\pi}{\omega}$, we obtain:

>>>>

% In array signal processing, propagating waves carry signals from the source to the array.
% Therefore, these signals are represented in the time-space domain. The space domain is represented by either the three dimensional Cartesian coordinates $\rBrace{x,y,z}$ or the three dimensional spherical coordinates $\rBrace{r,\phi,\theta}$, where $0 \leq \phi \leq 2\pi$, $0 \leq \theta \leq \pi$ are the azimuth and elevation angles, respectively. The time domain is represented by $t$.
An elementary physical phenomenon is the spatial dynamics of waves, referred to as wave-propagation.
The spatial location is noted by the Cartesian coordinates $\rBrace{x,y,z}$ or the three dimensional spherical coordinates $\rBrace{r,\phi,\theta}$, where $0 \leq \phi \leq 2\pi$, $0 \leq \theta \leq \pi$ are the azimuth and elevation angles, respectively. 
% Finally, let $f\rBrace{t,\vecnot{r}}$ denote the time-space representation of the input signal, where $\vecnot{r}$ is the radius-vector in the three-dimensional system. 
% The relations between the Cartesian and spherical coordinates are given in Fig.~\ref{fig_coordinates}.
Denoting $t$ as the time, the time-space representation of the a signal is $f\rBrace{t,\vecnot{r}}$ or , where the relations between the Cartesian and spherical coordinates are given in Fig.~\ref{fig_coordinates}.
% $f\rBrace{t,x,y,z}$ describes the signal impinging the microphone array. 
In a homogeneous, dispersion free and lossless medium the wave equation is:

@@

where $\nabla^{2}$ is the Laplacian operator and $c$ represents the wave's velocity in the medium.

@@

A possible solution to \eqref{eq_prlm_waveEq}, $f_p\rBrace{t,x,y,z}$, may be of a complex exponential form:

@@

Plugging \eqref{eq_prlm_waveEq_pSol} into \eqref{eq_prlm_waveEq} results in the \emph{monochromatic plane wave}:

@@

Using the plane wave notation emphasizes the fact that for a given time, $t_{0}$, all points on a plane given by $k_{x}x+k_{y}y+k_{z}z = constant$ are with the same wave value and the vector notation of \eqref{eq_prlm_waveEq}'s solution is

@@

where $\vecnot{k}\vecnot{x}=constant$ are planes of constant wave value.
The wave propagation can be described as the traveling of the planes, stating that small steps of both space $\delta{\vecnot{x}}$ and time $\delta{t}$ result in the same wave value i.e. $f\rBrace{t+\delta{t},\vecnot{x}+\delta{\vecnot{x}}} = f\rBrace{t,\vecnot{x}}$, which yields

@@

where $c>0$ is the wave velocity in the medium.
The \emph{wavelength} ($\lambda$) denote the distance the plane wave propagated during a single temporal period of $T=\frac{2\pi}{\omega}$.

@@

Using \eqref{eq_prlm_waveEq_stepsEq} with $\delta{t} = \frac{2\pi}{\omega}$, we obtain:

### 2020-12-25 Modules/sec_prlm_sensorArrays.tex
Before applying meeting#1's comments

Microphone arrays consist of sets of microphones positioned in a way to function as a directional acoustic antenna. 
Microphone arrays are utilized for filtering signals in a space-time field. 
Filtering is enabled by exploitation of the incoming signals spatial characteristics. 
A desirable spatial filtering, i.e., beamforming, should result in enhancement of signals of interest, originated in a specific direction, while forcing suppression of undesired signals originated in other directions.
Microphone arrays are utilized for solving many signal processing problems: dereverberation, localization of a single source, noise reduction and source separation \cite{cohen2004multichannel,habets2006dual,pavlidi2013real}.
\par 
Numerous factors must be taken into consideration when designing a microphone array configuration.
Initially, the geometry of the microphone array plays an important role in the formulation of the processing algorithms, as it forces fundamental constraints on the array’s operation. 
In most cases, the array geometry is the first consideration in array design due to practical and physical constraints of the design. 
Therefore, the degree of freedom in choosing the array geometry is limited. 
Nevertheless, in some other crucial problems such as noise reduction or source separation, the geometry of the array may have little importance. 
For instance, Uniform Linear Arrays (ULA) can only handle one source direction at a time resulting in uncertainties and possibly direction ambiguities.
Therefore, ULA configurations must be ruled out when handling several signal sources.
Other array geometries such as non-uniform linear arrays and circular arrays, have been studied in the field \cite{liu2008design,van2004optimum}.
This work is based on ULA and therefore we elaborate on this geometry.
A conventional beamformer \cite{van2004optimum} is composed of a ULA of microphones. 
That is, the microphones are positioned on an axis with a uniform spacing between the microphones.
The general expression of the microphone positioning in a ULA is given by:
\begin{equation}
x_{m}=m\cdot{d}, m=0, 1, \dots, M-1
\end{equation}
where $m$ denotes the microphone index; $x_{m}$ denotes the position of the microphone having index $m$, where $x_{m} = 0$ corresponds to the left-hand side of the array, i.e., the location where the microphone indexed by $m = 0$ is positioned; $d$ denotes the spacing between microphones, and $M$ is the array size, i.e., the number of microphones in the array, as demonstrated in Fig.~\ref{fig_ULA}.
\begin{figure}[h!]
    \begin{center}
        \begin{overpic}[width=0.5\linewidth, 
        %grid, 
        tics=10,trim=0 0 0 0]{./Media/fig_ULA.png}
            % \put (50, 62.5) {\footnotesize{$r=0$}}
        \end{overpic}
    \end{center}
     \caption{A uniform linear array of size M and spacing d.}
    \label{fig_ULA}
\end{figure}

>>>>

% Microphone arrays consist of sets of microphones positioned in a way to function as a directional acoustic antenna. 
% Microphone arrays are utilized for filtering signals in a space-time field. 
% Filtering is enabled by exploitation of the incoming signals spatial characteristics. 
% A desirable spatial filtering, i.e., beamforming, should result in enhancement of signals of interest, originated in a specific direction, while forcing suppression of undesired signals originated in other directions.
% Microphone arrays are utilized for solving many signal processing problems: dereverberation, localization of a single source, noise reduction and source separation \cite{cohen2004multichannel,habets2006dual,pavlidi2013real}.
Sensor arrays are sets of specifically positioned sensors, where every sensor's output is fused together with the rest, enabling the exploitation of spatial sampling.
Instead of mere temporal sampling, the user also receives spatial information from the relations between samples which arrived at the same temporal snapshot from all sensors. 
As in the temporal domain, where filters are specified with respect to the frequency domain, spatial filters are specified in the spatial frequency domain, i.e. DOA, where the equivalent of pass band are the DOAs of interest and the attenuated DOAs (nulls) are equivalent to the stop-band.
Sensor arrays are used in numerous applications, ranging from source localization, through communication, medical applications, astronomy etc.
\par 
% Numerous factors must be taken into consideration when designing a microphone array configuration.
% Initially, the geometry of the microphone array plays an important role in the formulation of the processing algorithms, as it forces fundamental constraints on the array’s operation. 
% In most cases, the array geometry is the first consideration in array design due to practical and physical constraints of the design. 
% Therefore, the degree of freedom in choosing the array geometry is limited. 
% Nevertheless, in some other crucial problems such as noise reduction or source separation, the geometry of the array may have little importance. 
% For instance, Uniform Linear Arrays (ULA) can only handle one source direction at a time resulting in uncertainties and possibly direction ambiguities.
% Therefore, ULA configurations must be ruled out when handling several signal sources.
% Other array geometries such as non-uniform linear arrays and circular arrays, have been studied in the field \cite{liu2008design,van2004optimum}.

When designing a sensor array, multiple consideration are to be taken into account, ranging from physical characteristics e.g. area, weight and carrying platform, through performance related considerations such as accuracy, spatial selectivity, signal to noise ratio and even the financial aspect - i.e. costs should be considered.
Obviously, the most fundamental array specification is its geometry, for it dictates the spatial (and mathematical) relations between simultaneous samples measured by the array's sensors, which also greatly influences the processing methods of the raw-data. 
This work is based on ULA and therefore we elaborate on this geometry.
A conventional beamformer \cite{van2004optimum} is considered to be ULA, i.e. uniformly spaced sensors, where the convention is that the ULA is place on the horizontal axis.
The sensors are positioned according to:
\begin{equation}
x_{m}=m\cdot{d}, m=0, 1, \dots, M-1
\end{equation}
where $m$ is the sensors index; $x_{m}$ denotes the position of the $m$'th sensor,$p_{0} = 0$ corresponds to the left-hand side of the array, $d$ denotes the spacing between sensors, and $M$ is the array size, as demonstrated in Fig.~\ref{fig_ULA}.
\begin{figure}[h!]
    \begin{center}
        \begin{overpic}[width=0.5\linewidth, 
        %grid, 
        tics=10,trim=0 0 0 0]{./Media/fig_ULA.png}
            % \put (50, 62.5) {\footnotesize{$r=0$}}
        \end{overpic}
    \end{center}
     \caption{A uniform linear array of size M and spacing d.}
    \label{fig_ULA}
\end{figure}

### 2020-12-25 Modules/sec_prlm_timeSpaceSig.tex
Before applying meeting#1's comments

The incoming signals are spatially sampled by the array microphones, then the samples are processed to attenuate signals from undesired directions and extract the signal from a desired direction. 
A spatial response of the microphone array is obtained via a beam (main-lobe) directed to the desired signal while nulls are directed to the undesired signals.
\begin{figure}[ht!]
    \begin{center}
        \begin{overpic}[width=0.5\linewidth, 
        %grid, 
        tics=10,trim=0 0 0 0]{./Media/fig_ULA_imping.png}
            % \put (50, 62.5) {\footnotesize{$r=0$}}
        \end{overpic}
    \end{center}
     \caption{A simple beamformer design.}
    \label{fig_ULA_imping}
\end{figure}
Fig.~\ref{fig_ULA_imping} illustrates a beamformer design based on a ULA of Fig.~\ref{fig_ULA}.
Let $\vecnot{f}\rBrace{t,\vecnot{x}}$ denote the set of signals sampled by the microphone array at time $t$, expressed by
\begin{equation}
\vecnot{f}\rBrace{t,\vecnot{x}} = \vBrace{f\rBrace{t,x_{0}},\dots,f\rBrace{t,x_{M-1}}}^{T},
\end{equation}
where $x_{m}$ denotes the microphone position, $t$ denotes the continuous-time variable and $^{T}$ denotes the transpose operation. 
The beamformer output $y\rBrace{t}$ is expressed by:
\begin{equation}
y\rBrace{t} = \sum_{m=0}^{M-1}{f\rBrace{t,x_{m}}\omega_{m}^{*}},
\end{equation}
where $\omega_{m}$ is a complex weight of the microphone having index $m$, as shown in Fig.~\ref{fig_ULA_imping} and $^{*}$ denotes complex conjugation. 
The simplest beamformer design architecture has uniform weights, $\omega_{m}=\frac{1}{M}, m=0,\dots,M-1$.
Let $f_{\omega}\rBrace{t,x} = e^{j\omega{t}}$ denote a plane wave propagating at angular frequency $\omega$, and let $\theta \in \vBrace{-\frac{\pi}{2},\frac{\pi}{2}}$ denote the direction of arrival (DOA) angle measured with respect to the broadside of the linear array, as shown in Fig.~\ref{fig_ULA_imping}.
The wave signals spatially sampled by the microphone array inputs are given by:
\begin{equation}
\label{eq_prlm_timeSpace_outputVec}
\vecnot{f}\rBrace{t,x} = \vBrace{f_{\omega}\rBrace{t-\tau_{0}}, \dots, f_{\omega}\rBrace{t-\tau_{M-1}}}^{T}, \tau_{m} = \frac{\sin{\theta}\cdot{}x_{m}}{c},
\end{equation}
where $\tau_{m}$ is the propagation delay for the incoming signal and $c$ is the wave's velocity in the medium.
A value of $\tau_{m}=0, \forall{m}$ implies a DOA of $\theta = 0$, i.e. a plane wave parallel to the array, propagating perpendicularly to the array.
Let $\kappa=\frac{\omega}{c}\sin{\theta}=\frac{2\pi}{\lambda}\sin{\theta}$ denote the wavenumber for plane waves in a locally homogeneous medium, where λ denotes the wavelength corresponding to the angular frequency $\vecnot{v}\rBrace{\kappa}$ denote the \emph{array manifold vector} \cite{van2004optimum}, featuring all of the spatial characteristics of the microphone array.
Based on \eqref{eq_prlm_timeSpace_outputVec}, and the definition of κ above, the manifold vector can be expressed as
\begin{equation}
\vecnot{v}\rBrace{\kappa}=\vBrace{e^{-j\kappa{}x_{0}},\dots,e^{-j\kappa{}x_{M-1}}}^{T}.
\end{equation}

>>>>

% The incoming signals are spatially sampled by the array microphones, then the samples are processed to attenuate signals from undesired directions and extract the signal from a desired direction. 
% A spatial response of the microphone array is obtained via a beam (main-lobe) directed to the desired signal while nulls are directed to the undesired signals.
As mentioned before, the measurements of the sensor array's elements are fused together, where spatial information resides in the relations between samples taken in the same time $t$.
A basic sensor array based entity is the beamformer, which is designed to enhance signals from certain directions while suppressing other signals from unwanted directions, heuristically forming a beam directed to the enhanced direction.  
\begin{figure}[ht!]
    \begin{center}
        \begin{overpic}[width=0.5\linewidth, 
        %grid, 
        tics=10,trim=0 0 0 0]{./Media/fig_ULA_imping.png}
            % \put (50, 62.5) {\footnotesize{$r=0$}}
        \end{overpic}
    \end{center}
     \caption{A simple beamformer design.}
    \label{fig_ULA_imping}
\end{figure}
Fig.~\ref{fig_ULA_imping} illustrates a beamformer design based on a ULA of Fig.~\ref{fig_ULA}.
% Let $\vecnot{f}\rBrace{t,\vecnot{x}}$ denote the set of signals sampled by the microphone array at time $t$, expressed by
A temporal snapshot measured by the array at time $t$ is
\begin{equation}
\vecnot{f}\rBrace{t,\vecnot{x}} = \vBrace{f\rBrace{t,x_{0}},\dots,f\rBrace{t,x_{M-1}}}^{T},
\end{equation}
where $x_{m}$ denotes the microphone position, $t$ denotes the continuous-time variable and $^{T}$ denotes the transpose operation. 
Assuming simple, weight-and-sum beamformer, the output ($y\rBrace{t}$) is expressed by:
\begin{equation}
y\rBrace{t} = \sum_{m=0}^{M-1}{f\rBrace{t,x_{m}}w_{m}^{*}},
\end{equation}
where $w_{m}$ is a complex weight used for the $m$'th sensor's output, as shown in Fig.~\ref{fig_ULA_imping} and $^{*}$ denotes complex conjugation. 
Consider $s\rBrace{t} = e^{j\omega{t}}$ as a plane wave propagating at angular frequency $w$, and $\theta \in \vBrace{-\frac{\pi}{2},\frac{\pi}{2}}$ as the direction of arrival (DOA) angle measured with respect to the broadside of the linear array, as shown in Fig.~\ref{fig_ULA_imping}.
The wave signals spatially sampled by the sensor array inputs are
\begin{equation}
\label{eq_prlm_timeSpace_outputVec}
\vecnot{f}\rBrace{t,x} = \vBrace{s\rBrace{t-\tau_{0}}, \dots, s\rBrace{t-\tau_{M-1}}}^{T}, \tau_{m} = \frac{\sin{\theta}\cdot{}x_{m}}{c},
\end{equation}
where $\tau_{m}$ is the propagation delay for the incoming signal and $c$ is the wave's velocity in the medium.
% A value of $\tau_{m}=0, \forall{m}$ implies a DOA of $\theta = 0$, i.e. a plane wave parallel to the array, propagating perpendicularly to the array.
Let $\kappa=\frac{\omega}{c}\sin{\theta}=\frac{2\pi}{\lambda}\sin{\theta}$ denote the wavenumber for plane waves in a locally homogeneous medium, where $\lambda$ denotes the wavelength corresponding to the angular frequency $\vecnot{d}\rBrace{\kappa}$ denote the \emph{steering vector}, featuring all of the array's spatial characteristics.
Based on \eqref{eq_prlm_timeSpace_outputVec}, and the definition of $\kappa$ above, the steering vector can be expressed as
\begin{equation}
\vecnot{d}\rBrace{\kappa}=\vBrace{e^{-j\kappa{}x_{0}},\dots,e^{-j\kappa{}x_{M-1}}}^{T}.
\end{equation}
